Evaluating the Capabilities of the Lattice Boltzmann Method for non-Newtonian and Free-surface Flow towards Applications in Wellbore Cementing

### 2017-03-03 response-to-reviewers.tex

\documentclass{article}

\begin{document}

\section{Response to reviewers}

Thank you very much for your assistance in reviewing this paper. We are sure that with your help, this has become a much better work. We hope that the attached edited copy of "Numerical Investigation of the Accuracy, Stability, and Efficiency of Lattice Boltzmann Methods in Simulating Non-Newtonian Flow" will answer any questions previously raised, and will be satisfactory for publication.

The following is a list of all comments recieved, and how they have been addressed in the updated version:

  \subsection{Reviewer 1}
	\begin{enumerate}
		\item \emph{In Fig. 2, the scale of the figure should be modified in order to avoid readers’ misleading. If the authors calculate only one case with L = 32 and H = 64, the computational domain becomes vertically long.}

@@

		The results are actually the same for all of the BGK variants.
		This is because for $n = 1.25$ the nonequilibrium entropy threshold for median filtering was not exceeded, and hence it was effectively the same as the unfiltered BGK scheme.
		Similarly, the relaxation time was within the bounds for the BGK-BRT scheme, which again resulted in the scheme to being effectively the same as the BGK collision scheme without bounds.
		The error norms for the MRT collision scheme was different, but the difference could not be seen in the precision of the results presented in the previous draft.
		The $n = 1.25$ error norms have been updated with higher precision so as to show the difference.

		\item \emph{The authors carried out the simulations with fixed lattice resolution. However, the convergence rate should be examined by calculating the benchmark problem (e.g., Poiseuille flow) with different lattice resolutions. 
		What order is the accuracy of the schemes used in the present study?}

		Considerations regarding grid resolution were added in the following locations:
		\begin{itemize}
			\item Table 2 at the end of Section 3.1.1 and the last paragraph of Section 3.1.1 address the significance of grid resolution for Bingham plastic Poiseuille flow.
			\item Table 4 at the end of Section 3.1.2 and the last paragraph of Section 3.1.2 address the significance of grid resolution for Power-law Poiseuille flow. 
			\item The third to last paragraph of Section 3.2.1 addresses the significance of grid resolution for the Bingham plastic lid-driven cavity flow.
		\end{itemize}
	\end{enumerate}

	\subsection{Reviewer 2}
	  The present work examines the single-relaxation-time (SRT) and the multiple-relaxation-time (MRT) lattice Boltzmann method (LBM) for Bingham plastic and power-law fluids. Constitutive relations in these fluids are typically stiff presenting substantial numerical challenges when the relations are incorporated into the LBM collision models. Entropic filtering as additional artificial dissipation and MRT collision are considered as the means to stabilize otherwise oscillatory or unstable SRT LBM. Optimal value for the median filtering was suggested and computational costs with these approaches are compared in detail. The conclusions are rather trivial; MRT is the most accurate and stable whereas its computational cost can be significantly large compared with SRT due to non-linear iteration. Filtering does stabilize SRT but the accuracy deteriorates particularly when the filter threshold is not optimally chosen. This work can be useful if readers are interested in choosing a suitable LB model for simulating non-Newtonian flows. However, it lacks novelty. Although MRT collision operator is shown to be the best choice, no detailed analysis on the values of optimum relaxation rates for various non-Newtonian flows.
	As for Poiseulle Flow part, SRT simulations deviate from the analytical solution in the core region (Figs.3,4) due to the difference in the constitutive relations used in LBM and the analytical solution. Thus, the reported errors in Table 1 include both constitutive relation error and truncation error. This does not seem fair. It would be helpful if the analytical solution or a benchmark solution from NS equations of Papanastasiou approximation is directly compared with LBM results. The authors attributed the nonphysical oscillations of SRT simulations to sharp gradients in the velocity near the walls, which is then again a result of a sharp gradient in the viscosity. However, from Figs.3&4, it is difficult to observe any sign of sharp gradients in the velocity near the walls. In Fig.4 for higher m, the oscillations occur at the center of the computational domain and they are not symmetric.

### 2017-03-03 response-to-reviewers.tex

\begin{document}

	\pagenumbering{gobble}

\section*{Response to reviewers}

Thank you very much for your assistance in reviewing this paper. We are sure that with your help, this has become a much better work. We hope that the attached edited copy of ``Numerical Investigation of the Accuracy, Stability, and Efficiency of Lattice Boltzmann Methods in Simulating Non-Newtonian Flow'' will answer any questions previously raised, and will be satisfactory for publication.

@@

		%The results are actually the same for all of the BGK variants.
		%This is because for $n = 1.25$ the nonequilibrium entropy threshold for median filtering was not exceeded, and hence it was effectively the same as the unfiltered BGK scheme.
		%Similarly, the relaxation time was within the bounds for the BGK-BRT scheme, which again resulted in the scheme to being effectively the same as the BGK collision scheme without bounds.
		%The error norms for the MRT collision scheme was different, but the difference could not be seen in the precision of the results presented in the previous draft.
		%The $n = 1.25$ error norms have been updated with higher precision so as to show the difference.

		The Table referenced in the above comment (previously Table 2), along with the power-law Poiseuille flow section, have been removed. We decided to remove the power-law Poiseuille flow section because we do not believe it added much value to the manuscript, and because we wanted to address grid resolution effects without lengthening the manuscript.

		\item \emph{The authors carried out the simulations with fixed lattice resolution. However, the convergence rate should be examined by calculating the benchmark problem (e.g., Poiseuille flow) with different lattice resolutions. 
		What order is the accuracy of the schemes used in the present study?}

		Considerations regarding grid resolution were added in the following locations:
		\begin{itemize}
			\item Table 2 at the end of Section 3.1 and the last paragraph of Section 3.1 address the significance of grid resolution for Bingham plastic Poiseuille flow.
			\item The second to last paragraph of Section 3.2.1 addresses the significance of grid resolution for the Bingham plastic lid-driven cavity flow.
		\end{itemize}

@@

		\item \emph{Several grammatical errors occur in the text. The authors should be checked again throughout the manuscript.}

		The manuscript was checked carefully for spelling and grammatical errors. We appreciate you pointing this out and we hope that no more grammatical errors persist.

	\end{enumerate}

	\subsection*{Reviewer 2}

	\begin{itemize}

	\item \emph{Although MRT collision operator is shown to be the best choice, no detailed analysis on the values of optimum relaxation rates for various non-Newtonian flows.}

	Thank you for the suggestion. Another collision scheme was added to the Bingham plastic Poiseuille flow study so that the choices for the free relaxation parameters of $s_1 = s_2 = s_4 = s_6 = 1.1$ and $s_1 = 1.1, s_2 = 1.0, s_4 = s_6 = 1.2$ could be compared. The error and computational cost for each choice of relaxation parameters are shown in Table 1 and Table 2, and the two collision schemes are compared in the discussions that follow each table.

	\item \emph{As for Poiseuille Flow part, SRT simulations deviate from the analytical solution in the core region (Figs.3,4) due to the difference in the constitutive relations used in LBM and the analytical solution. Thus, the reported errors in Table 1 include both constitutive relation error and truncation error. This does not seem fair. It would be helpful if the analytical solution or a benchmark solution from NS equations of Papanastasiou approximation is directly compared with LBM results.}

@@

	The comment below equation 28 (page 17) has been expanded to include the above reasoning.

	\item \emph{The authors attributed the nonphysical oscillations of SRT simulations to sharp gradients in the velocity near the walls, which is then again a result of a sharp gradient in the viscosity. However, from Figs.3\&4, it is difficult to observe any sign of sharp gradients in the velocity near the walls. In Fig.4 for higher m, the oscillations occur at the center of the computational domain and they are not symmetric.}

	We believe that the oscillations are more pronounced at the center because the velocity is prescribed at the wall, but as particle distributions advect from the wall toward the center overrelaxation can occur. It would be interesting to explore the exact cause of the oscillations more in future work.

	The oscillations are not symmetric because Fig. 4 is merely a snapshot in time. They tend to move back and forth across the channel in subsequent time steps such that, on average (in time), the oscillations are symmetric. A sentence has been added to the end of the first paragraph of page 19 in order to clarify the apparent asymmetry.

@@

	As mentioned in the responses to Reviewer 1, considerations regarding grid resolution were added in the following locations:
	\begin{itemize}
		\item Table 2 at the end of Section 3.1.1 and the last paragraph of Section 3.1.1 address the significance of grid resolution for Bingham plastic Poiseuille flow.
		\item The second to last paragraph of Section 3.2.1 addresses the significance of grid resolution for the Bingham plastic lid-driven cavity flow.
	\end{itemize}

	\end{itemize}

Thank you again for your help. Please let us know of any further improvements that can be made to this paper.

\vspace{1in}
\noindent Sincerely, \\
\indent John C. Brigham \\
\indent Associate Professor \\
\indent School of Engineering and Computing Sciences

### 2017-03-03 cover-letter-to-editor.tex
Updated Bingham Poiseuille section with MRT-1 and MRT-2
corrected in address in the cover letter
updated figure legends, MRT -> MRT-1
addressed MRT-1, MRT-2 comment in response

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}

\begin{document}

Fernando Pinho
Universidade do Porto, Porto, Portugal

Dear Prof. Pinho:

Thank you very much for your assistance with our article titled "Numerical Investigation of the Accuracy, Stability, and Efficiency of Lattice Boltzmann Methods in Simulating Non-Newtonian Flow" by Matthew Grasinger, Scott Overacker, and John C. Brigham for publication in the Journal of Non-Newtonian Fluid Mechanics. We are submitting to you the revised manuscript in which we have addressed the suggestions by the reviewers. In addition to the revised manuscript, we have included a document of our responses to the reviewers. Please let us know if any further modification or information is required.

@@

John C. Brigham
blah blah blah

\end{document}

>>>>

\documentclass{letter}
\signature{John C. Brigham \\
	\vspace{0.1in}
	Associate Professor \\
	School of Engineering \\ \hspace{0.1in} and Computing Sciences}
\address{South Road\\
	Durham\\
	DH1 3LE \\ United Kingdom}
\begin{document}
	\begin{letter}{Fernando Pinho \\ Universidade do Porto, Porto, Portugal}
		\opening{Dear Prof. Pinho,}
		Thank you very much for your assistance with our article titled ``Numerical Investigation of the Accuracy, Stability, and Efficiency of Lattice Boltzmann Methods in Simulating Non-Newtonian Flow'' by Matthew Grasinger, Scott Overacker, and John C. Brigham for publication in the Journal of Non-Newtonian Fluid Mechanics. We are submitting to you the revised manuscript in which we have addressed the suggestions by the reviewers. In addition to the revised manuscript, we have included a document of our responses to the reviewers. Please let us know if any further modification or information is required.

@@

		John C. Brigham\\
		South Road\\
		Durham\\
		DH1 3LE\\
		john.brigham@durham.ac.uk

		\vspace{1in}
		\closing{Sincerely,}
	\end{letter}
\end{document}
